---
title: "R Packages for SDS"
date: "September 7, 2016"
output: html_document
---

```{r}
library(knitr)
opts_chunk$set(eval = FALSE)
```



## Check existing library paths

```{r}
libs <- .libPaths()
libs
```

> Note: `admin` installs into `/usr/local/lib/R/site-library`

```{r}
sys.lib <- "/usr/local/lib/R/site-library"
```

## Currently installed packages

```{r}
pkgs <- as.data.frame(installed.packages(sys.lib))
pkgs.here <- as.character(pkgs$Package)
# these are the packages that are installed in the system library
pkgs.here
```

## Update currently installed packages

```{r}
# run this first -- should always be safe
update.packages(lib.loc = sys.lib)
```

## SDS packages requested

These packages should be installed in the system library for all users. Please update these **every semester**

```{r}
pkgs.desired <- c("mosaic", "dplyr", "rmarkdown", "knitr", "markdown", "formatR"
                  , "rstudio", "ggplot2", "tidyr", "stringr", "agricolae", "leaps"
                  , "Lahman", "twitteR", "latticeExtra", "RSQLite"
                  , "igraph", "Stat2Data", "lme4", "sp", "maptools", "rpart"
                  , "manipulate", "BiocInstaller", "mosaicData", "RcolorBrewer"
                  , "nycflights13", "devtools", "googlesheets", "readr"
                  , "mdsr", "etl", "macleish", "lubridate", "openintro"
                  , "tm", "jsonlite", "ggmap", "ggvis", "Rcpp", "plotly"
                  , "tigris", "rvest", "MASS", "XML", "leaflet"
                  , "readxl", "rmarkdown", "DBI", "ISLR", "broom")
```

## Install missing requested packages

```{r}
# run this -- should always be safe but may take a while
install.packages(setdiff(pkgs.desired, pkgs.here), lib = sys.lib)
```

## Non-standard installs

These packages may have external dependencies that require some finagling.

```{r}
pkgs.tricky <- c("RMySQL", "RPostgreSQL", "rgdal", "rgeos")
install.packages(setdiff(pkgs.tricky, pkgs.here), lib = sys.lib)
```

These packages are not on CRAN and must be installed from GtiHub.

```{r}
pkgs.github <- c("andrewpbray/oilabs", "rstudio/sparklyr", "omegahat/Sxslt")
lapply(pkgs.github, FUN = "devtools::install_github")
```

## Find multiply installed packages

Note that packages can be installed in multiple libraries!

```{r}
library(dplyr)
as.data.frame(installed.packages()) %>%
  group_by(Package) %>%
  summarise(N = n(), where = paste(LibPath, collapse = ", ")) %>%
  filter(N > 1) %>%
  print.data.frame()
```